Spatial Analysis

### 2021-07-28 Scripts/spatial_analisis.Rmd
Updated maps and port shapefile fom GFW

```{r setup, include=FALSE}
library(MyFunctions)
MyFunctions::my_lib(c("ggmap","sf","st","tidyverse","tools","readr","data.table","maps","viridis","wesanderson","knitr","kableExtra","plotly","ggrepel","ggsflabel","janitor","interp","rnaturalearth"))
```

@@

We use three different shapefiles for the continental U.S. land mass, the State waters of maine, new hampshire, massachusets, connecticut, rhode island, new york, new jersey, delaware, maryland, virginia, north carolina and the North East EEZ.

1.- **Land shapefile;** covers the US land territory for visualization. Data provided from the `map` package.

2.- **State waters;** covers the state waters of the NE US states. Data from [data.gov](https://catalog.data.gov/dataset/federal-and-state-waters).

@@

3.- **EEZ shapefile;** Used the *Sea Around Us* shapefle updated to June 2016.

```{r maps_sf, eval = T, echo = T, results = 'hide'}

# US State Map (land)

land_sf <- st_as_sf(map("state", plot = FALSE, fill = TRUE)) %>% 
  filter(ID %in% States) %>% 
  mutate(abrev = c("CO","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VI"))

# ggplot(us_map) +
  # geom_sf()

# US EEZ map

# path_world <- my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile")
path_world <- "~/OneDrive - UW-Madison/Spatial_data/World_EEZ_v11_20191118/"

# The File
fnam_world <- "eez_boundaries_v11.shp"

# Load it!

eez_sf <- st_read(dsn = path_world,
                        layer =file_path_sans_ext(fnam_world)) %>%
  # rename(eez_name = Name) %>%
  # st_transform(crs = 4326) %>% # 4326
  filter(TERRITORY1 == "United States",
         LENGTH_KM == 2734) #%>%
  # filter(str_detect(EEZ1, "United States")) %>%
  # st_simplify(preserveTopology = TRUE, dTolerance = 0.1) #0.1 for paper

# ggplot() +
#   geom_sf(data = World_eez_sf, aes(), color = "red") +
#   geom_sf(data = us_map, aes(), color = "blue")

# US state waters
# https://catalog.data.gov/dataset/federal-and-state-waters

# us_state_w <- rgdal::readOGR(dsn = "~/Downloads/FederalAndStateWaters/FederalAndStateWaters.gdb")
state_sf <-  st_read("~/OneDrive - UW-Madison/Spatial_Data/USFedAndStateW/FederalAndStateWaters.gdb") %>%
  janitor::clean_names() %>% 
  mutate(jurisdicti = str_to_lower(jurisdicti)) %>% 
  filter(jurisdicti %in% States) %>% 
  rename(state = jurisdicti) %>% 
  # st_transform(crs = 4326)  # for matching projections
  st_simplify(preserveTopology = TRUE, dTolerance = 10000) #0.1 for paper

grid_sf <-  st_read("/Volumes/Enterprise/Data/AcrossBoundaries/Data/Spatial/grid_sf/grid_eez_sf.shp")

unique(land_sf$ID)

ggplot() +
  geom_sf(data = eez_sf, aes()) +
  geom_sf(data = state_sf, aes(color = state), fill = NA) +
  scale_color_manual(values = state_pallet)+
  theme_dark()

@@

As a first step we need to divide the NE US EEZ among the different states. For that we expanded state waters up to the 200 nautical miles to then estimate the percentage that each expanded-state-waters occupied. Note that in all cases these areas overlapped and percentages accounted for it. We did this by following these steps:

- 1. Expand state waters using a buffer

@@

@@

>>>>

```{r setup, include=FALSE}
library(MyFunctions)
MyFunctions::my_lib(c("ggmap","sf","st","tidyverse","tools","readr","data.table","maps","viridis","wesanderson","knitr","kableExtra","plotly","ggrepel","ggsflabel","janitor","interp","rnaturalearth","readxl"))

# Fix new updats of sf package
sf::sf_use_s2(use_s2 = FALSE)

@@

We use three different shapefiles for the continental U.S. land mass, the State waters of maine, new hampshire, massachusets, connecticut, rhode island, new york, new jersey, delaware, maryland, virginia, north carolina and the North East EEZ. Moreover, we use the NOAA dataset of landings per port.

##### 1.- Land shapefile

Covers the US land territory for visualization. Data provided from the `map` package.

```{r land_sf, eval = T, echo = T, results = 'hide'}

# -------------- #
# US State Map (land)
# -------------- #

land_sf <- st_as_sf(map("state", plot = FALSE, fill = TRUE)) %>% 
  filter(ID %in% States) %>% 
  mutate(abrev = c("CO","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VI"),
         state = str_to_sentence(ID)
         )

# Visual exploration (o.k! )
ggplot(land_sf) +
  geom_sf(aes(fill = state)) +
  scale_fill_manual(values = state_pallet) +
  theme_dark()

#####  2.- EEZ shapefile

Used the *Sea Around Us* shapefle updated to June 2016.

```{r eez_sf, eval = T, echo = T}

# US EEZ map
path_world <- my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile", name = "SAUEEZ_July2015.shp")

# Load it!
eez_sf <- st_read(path_world) %>%
  rename(eez_name = Name) %>%
  st_transform(crs = 4326) %>% # 4326
  filter(eez_name == "USA (East Coast)") %>%
  st_simplify(preserveTopology = TRUE, dTolerance = 0.1) #0.1 for paper

# Visual exploration (o.k! )
ggplot(eez_sf) +
  geom_sf() +
  theme_dark()

#####  3.-  State waters

Covers the state waters of the NE US states. Data from [data.gov](https://catalog.data.gov/dataset/federal-and-state-waters).

@@

```{r state_sf, eval = T, echo = T}

# us_state_file <- my_path("G", extra_path = "Spatial/USFedAndStateW/FederalAndStateWaters.gdb")

state_sf <-  st_read("~/OneDrive - UW-Madison/Spatial_Data/USFedAndStateW/FederalAndStateWaters.gdb") %>%
  janitor::clean_names() %>% 
  mutate(jurisdicti = str_to_lower(jurisdicti)) %>% 
  filter(jurisdicti %in% States) %>% 
  mutate(state = str_to_sentence(jurisdicti)) %>% 
  st_transform(crs = 4326) %>%   # for matching projections
  st_simplify(preserveTopology = TRUE, dTolerance = 10000) #0.1 for paper

# Visual exploration (o.k! )
ggplot(state_sf) +
  geom_sf(aes(fill = state)) +
  scale_fill_manual(values = state_pallet) +
  theme_dark()

##### 5.- Fishing ports

We used [Global Fishing Watch](https://globalfishingwatch.org/datasets-and-code-anchorages/)'s database on 
Anchorages, Ports and Voyages.

*The Global Fishing Watch anchorages dataset is a global database of anchorage locations where vessels congregate. The dataset contains over 160,000 individual anchorage locations, which are further grouped into nearly 32,000 ports when applicable.*

```{r ports_sf, eval = T, echo = T}

# ports_data <- my_path("D","NOAA", name = "SSmith_2015-2021 Landings_JUN 2021.xlsx", read = T) %>% 
#   slice(-50) %>% # remove uneccessary row
#

gfw_data <- my_path("D","Spatial",name = "gfw_portdata.csv", read = T) %>% 
  filter(iso3 == "USA")

# Visualize data (o.k!)
# ggplot(gfw_data) +
#   geom_point(
#     aes(
#     x = lon,
#     y = lat
#   )
#   )

# Convert data to sf for mergig
wpi_sf <- st_as_sf(wpi_data,
             coords = c("longitude", "latitude"),
             crs = 4326)

# Merge data with state waters to,locate ports 
port_sf <- st_join(wpi_sf,state_sf) %>% 
  filter(!is.na(jurisdicti)) # remove points outside stat waters

ggplot(port_sf)  +
  geom_sf(aes(color = jurisdicti)) +
  scale_color_manual(values = state_pallet) +
  theme_dark()

This is how all the shapefiles look like together

```{r all_sf, eval = T, echo = T}
# Visual exploration of all together (o.k! )
ggplot() +
  geom_sf(data = eez_sf, aes()) +
  geom_sf(data = land_sf, aes(fill = state), alpha = 0.3) +
  geom_sf(data = state_sf, aes(color = state), fill = NA) +
  geom_sf(data = port_sf, aes(color = state)) +
  scale_color_manual(values = state_pallet) +
  scale_fill_manual(values = state_pallet) +
  theme_dark()

@@

As a first step we need to divide the NE US EEZ among the different states. For that, we expanded a buffer-polygon to the 200 nautical miles to then estimate the percentage that each expanded polygon occupied. For the initial buffer point we used two approaches; state waters and US ports. Note that in all cases these areas overlapped and percentages accounted for it. We did this by following these steps:

- 1. Expand state waters / US ports using a buffer

@@

##### 1.1 State Waters

##### 1.1 US. Ports

@@

# Recycling code
## Sarah's data

# This dataset includes all landings from 2015-July 2021. It includes the species landed, the port, and the Hull ID and vessel name for black sea bass and summer flounder.

# *License, from Sarah Smith: This is confidential data, so please do not share this file with others. You may need to contact NOAA directly just to request permission. My contact has been Alison Ferguson (Alison.Ferguson@noaa.gov) in their data department.*

### SRAH LANDING DATA
# read_excel_allsheets <- function(filename, tibble = TRUE) {
#     # I prefer straight data.frames
#     # but if you like tidyverse tibbles (the default with read_excel)
#     # then just pass tibble = TRUE
#     sheets <- readxl::excel_sheets(filename)[2:8]
#     x <- lapply(sheets, function(X) readxl::read_excel(filename, sheet = X))
#     if(!tibble) x <- lapply(x, as.data.frame)
#     names(x) <- sheets
#     x
# }
# # Get data path
# ports_path <- my_path("D","NOAA", name = "SSmith_2015-2021 Landings_JUN 2021.xlsx")
# 
# ports_data <- bind_rows(read_excel_allsheets(ports_path)) %>% 
#   clean_names() %>% 
#   mutate_if(is.character,
#               ~ str_to_lower(.))
